---
title: "Linear Modeling Notebook"
author: "Julian Freedberg"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```



## Linear Modeling

This code can be adjusted to run for all routes and calculates the R squared and MSE of all the routes and the number of folds used to test the data. We will be using a 80/20 testing training split for this data.  We can adjust the number of folds and the testing trianing split as needed in the code. It also generates some diagnostic charts that we used to evaluate the strength of our results.
```{r}
library(tidyverse)
library(vroom)
vp_clean = vroom("data/route_800.csv", delim = ",")

set.seed(12345)

unique_routes = unique(vp_clean$route_id)

r2_train = c()
r2_test = c()
route = c()
mse_test = c()
fold = c()
test_df = tibble()
trained = F
n_folds = 1

len = length(unique_routes)
for(i in 1:(length(unique_routes))){

  r = unique_routes[i]
    print(paste("Route:", r))
    
  #filtering rows that are far from the route shape
  vp <- vp_clean %>% filter(route_id == r) 
  
  #randomizing order of dataset 
  vp = vp %>% 
    sample_n(dim(vp)[1]) 
  
  #setting breaks for cross validation
  n = dim(vp)[1]
  if(n < 200){
    next
  }
  
  tbreak = c(0, floor(n * 0.2), floor(n * 0.4), floor(n * 0.6), floor(n * 0.6), n)
  
  # 5 fold cross validation
  for(j in 1:n_folds){
    if(trained == T){
      next
    }
    #splitting into testing and training dataset
    test = vp %>% slice(tbreak[j]:tbreak[j + 1])
    test_stops = test %>% select(stop_name) %>% distinct()

    train = vp %>% anti_join(test %>% select(id), by = c("id"))
    
    train_stops= train  %>% select(stop_name) %>% distinct()
    if(dim(test_stops)[1] != dim(test_stops %>% inner_join(train_stops, by = "stop_name"))[1]){
      next
    }
    else{
      trained = T
    }
    
   
   if(length(unique(train$time_of_day)) == 1){
     train$time_of_day = sample(c(train$time_of_day[1],"2"), dim(train)[1], replace = T)
   }
    trained = F
   
    reg = lm(delay ~ dist_to_stop * sched_speed  * time_of_day + stop_name + as.factor(weekday), data = train)
    
    #assinging variable
    assign(paste("reg", r, sep = "_"),reg)
    
    yhat = predict(reg, test)
    
    #adding predictions to test dataset
    test <- test %>% 
      mutate(yhat = yhat) 
    test_df <- bind_rows(test_df, test)
    
    yhat_train = predict(reg, train)
    
    
    #storing training error
    r2_train[j + (i - 1) * n_folds] <- cor(train$delay, yhat_train)^2
    
    print(r)
    print("Train RMSE")
    print(sqrt(sum((train$delay - yhat_train)^2)/dim(train)[1]))
    print("Train Accuracy")
    print(sum((floor(train$delay/60) - floor(yhat_train/60)) == 0)/dim(train)[1])
    print("Test Accuracy")
    print(sum((floor(test$delay/60) - floor(yhat/60)) == 0)/dim(test)[1])
    
    #storing rotue
    route[j + (i - 1) * n_folds] <- r
    mse_test[j + (i - 1) * n_folds] <- test %>% mutate(se = (yhat - delay) ^ 2 ) %>% 
      summarize(mse = mean(se)) %>% pull()
    r2_test[j + (i - 1) * n_folds] <- cor(test$delay, test$yhat)^2
    fold[j + (i - 1) * n_folds] = j
    
  }

}
```

Modeling results
```{r}
model_results = tibble(route, mse_test, r2_test, r2_train, fold)


knitr::kable(model_results)

test %>% 
  mutate(diff = floor((delay - yhat)/60)) %>% 
  ggplot(aes(x = diff)) + 
  geom_histogram(color = "white", breaks = seq(-5,13,1) - 0.011) + 
  labs(x = "Delay - Predicted Arrival (Minutes)",
       title = "Route 800: Distribution of Residuals",
       y = "Number of Observations") +
  scale_x_continuous(breaks = seq(-5,13,1)) 


vp_clean %>% 
  mutate(minute_to_arrival = floor(delay/60)) %>% 
  ggplot(aes(x = minute_to_arrival)) + 
  geom_histogram(breaks = c(0,1,2,3,4,5,6,7,8,9,10, 11, 12, 13, 14,15), color = "white") +
  labs(x = "Minutes to Arrival",
       y = "Number of Observations",
       title = "Distribution of Minutes to Arrival to Next Stop - Route 800")

  
```